---
title: "Stats 506 F20 Final Project"
author: "Aravind Mantravadi"
date: "12/11/2020"
output: 
  html_document:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

# 79: -------------------------------------------------------------------------

```

# Introduction

This project answers the question "do people in the US get their macronutrient
intake primarily at home or from elsewhere?" Eating outside is often assumed to 
be unhealthy, while home-cooked meals are supposed to provide a more nutritious 
and balanced diet. What does this tell us about the macronutrient split based 
on their source?

To analyze this, I used the NHANES 2017-18 Nutrition Dataset and the 2017-18 
Demographics Dataset and gathered the variables that contained survey weights, 
macronutrient intake and the source of the nutrient. There are 28 different 
sources of food and I manually classified them into "home-cooked" and 
"outside". For more detail about the classification, refer to the 
**Data and Methods** section. 

Finally, I found the mean of each macronutrient along with a 95% confidence 
interval, split by the source of food. This gave me a way to compare the 
average macronutrient intake by source.


# Data and Methods

I used the
[NHANES 2017-18 Nutrition Dataset](https://wwwn.cdc.gov/nchs/nhanes/search/datapage.aspx?Component=Dietary&CycleBeginYear=2017) and the 
[NHANES 2017-18 Demographic Dataset](https://wwwn.cdc.gov/nchs/nhanes/Search/DataPage.aspx?Component=Demographics&CycleBeginYear=2017)


For dataset manipulation, I used the **dplyr package** in tidyverse 
collection of packages. I merged the two datasets and kept only the 
variables of interest -- the survey weights, macronutrients, total 
food intake and source of food. Then, I manipulated the source of food 
as mentioned in the Introduction section and then used the functions 
provided by the **survey package** to find the average of each 
macronutrient for each source of food. The estimates are given with a 
95% confidence interval and they take into account the survey weights.


The results were compared visually using tables and charts, and they 
consist of a frequency table of the data, a table 
of means for each macronutrient, and a bar graph showing the food intake 
from the different sources. The code for this analysis can be found [here](https://github.com/aravind1338/506F20FinalProject).


## Categorizing food source

I manually categorized the food sources listed in the variable "DR1FS". I chose 
to place codes 1, 10, 11, 12, 18, 19, 20 under "home-cooked" and the rest as 
"outside". You can find detailed information about the codes [here](https://wwwn.cdc.gov/Nchs/Nhanes/2017-2018/DR1IFF_J.htm#DR1FS)



# Results

```{r results, echo=FALSE, fig.width=15, fig.height=7}

source('final_proj.R')

summary_stats = dataset %>% count(Source)
knitr::kable(summary_stats, format = "html", 
             col.names = c("Source", "Frequency"), 
             caption = "Table 1: Frequency distribution of data by source") %>%
  kableExtra::kable_styling('striped', full_width = TRUE)

barplot(food_weight$food_weight, names.arg = c("Home-cooked", "Outside"), 
        col = "darkred", main = "Figure 1: Average gram weight of food 
        consumed in a day", 
        xlab = "Source", ylab = "Weight (grams)", 
        ylim = range(pretty(c(0, food_weight$food_weight))))


knitr::kable(macronutrients, format = "html", 
caption = "Table 2: Average macronutrient intake by source (grams). Format: 
estimate (lower, upper)") %>%
  kableExtra::kable_styling('striped', full_width = TRUE)
```

The results in the tables and figures above illuminate quite a few things.

1. From Table 2, we can see that people in the US **primarily** get their 
macronutrient intake from home-cooked meals rather than outside sources. 
But, one interesting exception is the split of fat consumed between 
home-cooked and outside sources.

2. Table 2 also shows us that the only macronutrient that people in the US 
get from outside sources more than from home-cooked meals is fats.

3. We can also see from Figure 1 that although there is more food consumed 
from outside sources than home-cooked ones (on average), the difference 
isn't extremely large. 


# Discussion

From the results above, we can conclude a couple of things:

1. Result 2 would support the fact that the majority of food that is not 
home-cooked is full of unhealthy trans fats.

2. From result 3, we see that despite the difference, fats are the exception. 
This might be an indicator of the imbalance in nutrients in outside food.


## Strengths

* The analysis takes into account survey weights, which enables us to 
expand the question to "all people in the US" rather than just the sample 
population.

* The intuitive understanding of the composition of food from outside 
sources is confirmed by the results, potentially indicating that how the 
sources are classified is at the very least valid.

## Weaknesses

* The classification of sources was subjective, based on my judgement of 
the descriptions in the NHANES documentation. A different but still valid 
interpretation could lead to different results.

* Potential sources of confounding, like whether the data is representative 
of the population (even with the survey weights)


