---
title: "Stats 506 F20 Final Project"
author: "Aravind Mantravadi"
date: "12/9/2020"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

# 79: -------------------------------------------------------------------------

```

# Introduction

This project answers the question "do people in the US get their macronutrient
intake primarily at home or from elsewhere?" Eating outside is often assumed to 
be unhealthy, while home-cooked meals are supposed to provide a more nutritious 
and balanced diet. What does this tell us about the macronutrient split based 
on their source?

To analyze this, I used the NHANES 2017-18 Nutrition Dataset and the 2017-18 
Demographics Dataset and gathered the variables that contained survey weights 
macronutrient intake and the source of the nutrient. There are 28 different 
sources of food and I manually classified them into "home-cooked" and 
"outside". For more detail about the classification, refer to the 
**Data/Methods** section. 

Finally, I found the mean of each macronutrient along with a confidence 
interval, split by the source of food. This gave me a way to compare the 
average macronutrient intake by source.


# Data/Methods

I used the
[NHANES 2017-18 Nutrition Dataset](https://wwwn.cdc.gov/nchs/nhanes/search/datapage.aspx?Component=Dietary&CycleBeginYear=2017) and the 
[NHANES 2017-18 Demographic Dataset](https://wwwn.cdc.gov/nchs/nhanes/Search/DataPage.aspx?Component=Demographics&CycleBeginYear=2017)


For dataset manipulation, I used the **dplyr package** in tidyverse 
collection of packages. I merged the two datasets and kept only the 
variables of interest. Then, I manipultaed the source of food as 
mentioned in the Introduction section and then used the functions 
provided by the **survey package**, specifically **svydesign**, 
**svyby** and **svymean** to conduct an analysis of the dataset. 
The code can be found [here](https://github.com/aravind1338/506F20FinalProject).


TODO: Explanation for home-cooked and outside

# Results

```{r results, echo=FALSE, fig.width=15, fig.height=7}

source('final_proj.R')

barplot(food_weight$food_weight, names.arg = c("Home-cooked", "Outside"), 
        col = "darkred", main = "Figure 1: Gram weight of food consumed", 
        xlab = "Source", ylab = "Weight (grams)", 
        ylim = range(pretty(c(0, food_weight$food_weight))))

knitr::kable(macronutrients, format = "html", 
caption = "Figure 2: Macronutrient intake by source (grams). Format: 
estimate (lower, upper)") %>%
  kableExtra::kable_styling('striped', full_width = TRUE) 
```

The results in the table above illuminate quite a few things: TODO


# Conclusion/Discussion

TODO


# TODO

* Need to add an explanation for how I divided the codes 
into "home-cooked" and "outside"

* Need to expand upon the results and add another graphical/tabular element

* Need to write up the conclusion in detail according to the project specifications


